Fast Automatic Bayesian Cubature Using Lattice Sampling

### 2019-07-18 BayesianCubaturePaper2018 - Rev1b.tex
Updates from Overleaf
revising

\newcommand{\hatvy}{\hat{\bm{y}}}
\newcommand{\haty}{\hat{y}}
\newcommand{\tvy}{\tilde{\bm{y}}}
\newcommand{\ty}{\tilde{y}}
\newcommand{\vzero}{\bm{0}}
\newcommand{\vone}{\bm{1}}
\newcommand{\tvone}{\tilde{\bm{1}}}
\newcommand{\mA}{\mathsf{A}}
\newcommand{\mC}{\mathsf{C}}
\newcommand{\rmC}{\mathring{\mathsf{C}}}
\newcommand{\mCtheta}{{\mathsf{C}_{\vtheta}}}
%\newcommand{\mCthetaInv}{{\mathsf{C}^{-1}_{\vtheta}}}
%\newcommand{\mCthetaMLE}{{\mathsf{C}_{\vthetaMLE}}}
%\newcommand{\mCthetaInvMLE}{{\mathsf{C}^{-1}_{\vthetaMLE}}}
\newcommand{\mCInv}{\mathsf{C}^{-1}}
\newcommand{\cov}{{\textup{cov}}}
\newcommand{\var}{{\textup{var}}}

@@

%\allowdisplaybreaks[4]
\begin{document}
\setlength\abovedisplayskip{1pt}
\setlength{\belowdisplayskip}{1pt}

@@

The primary contribution of this article is to demonstrate how the choice of a family of covariance kernels that match the low discrepancy sampling nodes facilitates fast computation of the cubature and the data-driven stopping criterion.  Our Bayesian cubature requires a computational cost of
\begin{equation} \label{eqn:OuralgoCost}
    \Order\bigl(N_{\textup{opt}}[n \$(f) + n\$(C) + n \log(n)] \bigr),
\end{equation} 
where $\$(f)$ is the cost of one integrand value, $\$(C)$ is the cost of a single covariance kernel value,  $\Order(n \log(n))$ is the cost of a fast Fourier transform, and $N_{\textup{opt}}$ is an upper bound on the number of optimization steps required to choose the hyperparameters. If function evaluation is expensive, such as coming from computationally intensive simulation, or $\$(f) = \Order(d)$ for large $d$, then $\$(f)$ might be similar in magnitude to $\log(n)$ in practice.  Typically, $\$(C) = \Order(d)$.  Note that the $\Order(n \log(n))$ contribution is $d$ independent.

By contrast to our fast algorithm, the typical computational cost for Bayesian cubature is
\begin{equation} \label{eqn:TheiralgoCost}
    \Order\bigl(N_{\textup{opt}}[n \$(f) + n^2\$(C) + n^3] \bigr),
\end{equation} 
which is explained in Section \ref{sec:bayes_cubature_algo}. Note that the work beyond evaluating the integrand in \eqref{eqn:TheiralgoCost} is much larger than that in \eqref{eqn:OuralgoCost}.

@@

Section \ref{sec:BC} explains the Bayesian approach to estimate the posterior cubature error and defines our automatic Bayesian cubature. Although much of this material is known, it is included for completeness.  We end Section \ref{sec:BC}  by demonstrating why Bayesian cubature is typically computationally expensive.
Section \ref{sec:fast_BC}  introduces the concept of covariance kernels that match the nodes and expedite the computations required by our automatic Bayesian cubature. 
Section \ref{sec:shift_invariant_kernel} implements this concept for shift invariant kernels and rank-1 lattice nodes.  It also describes how to avoid cancellation error for kernels of product form.  Numerical examples are provided in Section \ref{sec:NumExp} to demonstrate our new algorithm.  We conclude with a brief discussion.

@@

\begin{theorem} \label{thm:param} There are at least three approaches to estimating or integrating out the hyperparameters defining the Gaussian process from which the integrand is drawn: empirical Bayes, full Bayes, and generalized cross validation.  Under these three approaches we have the following:
\begin{allowdisplaybreaks}
\begin{align}
    \label{eqn_m_MLE}
m_\MLE &= \frac{\vone^T \mCInv_\vtheta \vy }{ \vone^T \mCInv_\vtheta \vone}, \qquad
m_{\textup{GCV}} = \frac{\vone^T \mC_\vtheta^{-2} \vy}{\vone^T \mC_\vtheta^{-2} \vone}, \\
\label{eqn_s2_MLE}
s^2_{\MLE} 
&= 
\frac{1}{n}
\vy^T 
\left[ \mCInv_\vtheta - 
\frac{ \mCInv_\vtheta \vone \vone^T \mCInv_\vtheta }{\vone^T\mCInv_\vtheta \vone}
\right] \vy, \\
\nonumber
\hsigma_{\textup{full}}^2 
& = \frac{1}{n-1}
\vy^T\left[ \mC_\vtheta^{-1} 
- \frac{ \mC_\vtheta^{-1} \vone\vone^T \mC_\vtheta^{-1}}{\vone^T \mC_\vtheta^{-1} \vone}  \right]\vy
\\ 
& \times  \left[\frac{(1 - \vc_{\vtheta}^T \mC_\vtheta^{-1} \vone)^2}{\vone^T \mC_\vtheta^{-1} \vone} + (c_{0\vtheta}  -\vc_{\vtheta} ^T \mC_\vtheta^{-1} \vc_{\vtheta}) \right], \\
\nonumber
 s^2_{\textup{GCV}} & = \vy^T \left[\mC_\vtheta^{-2} - \frac{\mC_\vtheta^{-2} \vone \vone^T \mC_\vtheta^{-2}}{\vone^T \mC_\vtheta^{-2} \vone}  \right] \vy  \left[ \trace(\mC_\vtheta^{-1}) \right]^{-1}, \\
\nonumber
\vtheta_\MLE
&= \argmin_{\vtheta} \biggl \{
\log\left(\vy^T 
\left[ \mC_\vtheta^{-1} - 
\frac{ \mCInv_\vtheta \vone \vone^T \mCInv_\vtheta }{\vone^T\mCInv_\vtheta \vone}
\right] \vy 
\right)  \\
\label{eqn:thetaMLE}
 & \qquad +  \frac{1}{n} \log(\det(\mC_\vtheta))
\biggr \}, \\
\nonumber
\vtheta_{\textup{GCV}} &= \argmin_\vtheta \biggl\{\log \left(  \vy^T \left[\mC^{-2}_\vtheta - \frac{\mC^{-2}_\vtheta \vone \vone^T \mC^{-2}_\vtheta}{\vone^T \mC^{-2}_\vtheta \vone}  \right] \vy \right)  \\
\label{vthetaGCV}
& - \log \left ( \trace(\mC^{-2}_\vtheta) \right ) \biggr\}, \\
\hmu_\MLE  &= \hmu_\full =
\left(
\frac{ (1 - \vone^T  \mCInv_\vtheta\vc_{\vtheta} )  \vone }{ \vone^T \mCInv_\vtheta \vone}   +  \vc_{\vtheta} 
\right)^T  \mCInv_\vtheta \vy, \\
\label{eqn:muCV}
\widehat{\mu}_{\textup{GCV}}
& = \left(\frac{(1 - \vone^T  \mC_\vtheta^{-1}\vc_{\vtheta}) \mC_\vtheta^{-1} \vone}{\vone^T \mC_\vtheta^{-2} \vone} + \vc_{\vtheta} \right)^T \mC_\vtheta^{-1} \vy, \\
\nonumber 
\err_{\mathsf{x}} & = 2.58 s_{\mathsf{x}} \sqrt{c_{0\vtheta} - \vc_{\vtheta}^T\mC_\vtheta^{-1}\vc_{\vtheta} }, \quad \mathsf{x} \in \{\MLE, \GCV\}, \\
\nonumber 
\err_{\textup{full}} 
& = t_{n-1,0.995} \hsigma_{\textup{full}} > \err_\MLE,
\end{align}
\end{allowdisplaybreaks}
\begin{multline}
\label{eqn_prob_CI}
\mathbb{P}_f \left[
|\mu-\hmu_{\mathsf{x}}| \leq \err_{\mathsf{x}} \right]  = 99\%, \\ \mathsf{x} \in \{\MLE, \full, \GCV\}.
\end{multline}
Here $t_{n-1,0.995}$ denotes the $99.5$ percentile of a standard Student's $t$-distribution with $n-1$ degrees of freedom.
\end{theorem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Empirical Bayes}  \label{sec:MLE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The empirical Bayes approach estimates the parameters, $m$, $s$, and $\vtheta$ via maximum likelihood estimation (MLE).  The log-likelihood function of the parameters given the function data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv_\vtheta(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det\, \mC_\vtheta) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$ and then with respect to $s$ yields the values given in Theorem \ref{thm:param}.
To obtain $\vtheta_\MLE$, we substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$ and then numerically optimize:
\begin{align}
\nonumber
\vtheta_\MLE
&= \argmin_{\vtheta} \biggl \{
\log\left(\vy^T 
\left[ \mCInv_\vtheta - 
\frac{ \mCInv_\vtheta \vone \vone^T \mCInv_\vtheta }{\vone^T\mCInv_\vtheta \vone}
\right] \vy 
\right)  \\
\label{eqn:thetaMLE}
 & \qquad +  \frac{1}{n} \log(\det(\mC_\vtheta))
\biggr \}.
\end{align}
The empirical Bayes estimate of $\vtheta$ balances minimizing the covariance scale factor, $s^2_{\MLE}$, against minimizing  $\det(\mC_\vtheta)$. 
Under these estimates of the parameters, the cubature \eqref{eqn:BayesCub} and the credible interval \eqref{eqn_prob_confidence_interval} are explicitly written as in Theorem \ref{thm:param}.
The quantities $c_{0\vtheta}$, $\vc_\vtheta$, and $\mC_\vtheta$ are assumed implicitly to be based on $\vtheta = \vtheta_\MLE$.

@@

        \State\label{LoopStart}Generate $\{ \vx_i\}_{i=n' + 1}^{n}$ and sample $\{f(\vx_i)\}_{i=n'+1}^{n}$
        \State Compute $\vtheta$ by \eqref{eqn:thetaMLE} or \eqref{vthetaGCV}
        \State Compute $\err_{\textup{CI}}$  according to \eqref{eqn:errMLE}, \eqref{FJH:eq:errFull}, or \eqref{GCVerr}

@@

        \State Update sample size to compute $\hmu$, $n \gets n'$
        \State Compute $\hmu$, the approximate integral,   according to \eqref{eqn:cubMLE} or \eqref{eqn:muCV}
      \State \Return $\hmu, \; n$  and $\err_{\textup{CI}}$
  \end{algorithmic}
\end{algorithm}

\FJHNote{Add something about the cost.}

@@

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1)
\\
\alpha_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\\
\beta_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the cumulative standard normal distribution function.  Then, $\mu = \int_{[0,1]^{d'-1}} f_{\text{Genz}}(\vx) \, \dvx$. As we can see, this approach transforms $d'$ integral into a $d=d'-1$ integral.

@@

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol' points \cite{DicEtal14a,DicPil10a}.  The results are for 400 randomly chosen $\varepsilon$ in the interval $[10^{-5}, 10^{-2}]$ as shown in \figref{fig:MVN_Metern_d2b2}. In each run, the nodes are randomly scrambled.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

The generic automatic Bayesian cubature algorithm described in the previous section requires $\Order(n^3)$ operations to estimate $\vtheta$, compute the credible interval width, and compute the cubature. Now we explain how to speed up the calculations. A key is to choose covariance kernels that match the nodes, $\{\vx_i\}_{i=1}^n$, so that the vector-matrix operations required by Bayesian cubature can be accomplished using fast transforms at a cost of $\Order(n \log(n))$.

@@

Under assumptions \eqref{fastcompAssump} the eigenvalues may be identified as the fast transform of the first column of $\mC$:
\begin{align}
\nonumber
\vlambda 
& = \begin{pmatrix}
\lambda_1 \\ \vdots \\ \lambda_n
\end{pmatrix} = \mLambda \vone = \mLambda \vv_1^* 
= \underbrace{\left( \frac 1n \mV^H  \mV \right) }_{\mathsf{I}} \mLambda \vv_1^* \\
&= \mV^H \left( \frac 1n \mV \mLambda \vv_1^* \right)
= \mV^H \vC_1 =  \widetilde{\vC}_1,
\label{eqn:fast_transform_to_eigvalues}
\end{align}
Where $\mathsf{I}$ is the identity matrix.
Also note that the fast transform of $\vone$ has a simple form
\begin{align*} 
\widetilde{\vone}
& = \mV^H \vone = \mV^H \vV_1 = 
\left(n, 0, \dots, 0 \right)^T.
% \begin{pmatrix}n  \\ 0 \\ \vdots \\ 0 \end{pmatrix}.
\label{eqn:fast_transform_one}
\end{align*}

@@

The preceding sections lay out an automatic Bayesian cubature algorithm whose computational cost is only $\Order(n \log(n))$ if $n$ function values are used.  However, this algorithm relies on covariance kernel functions, $C$ and node sets, $\{\vx_i\}_{i=1}^n$ that satisfy assumptions \eqref{fastcompAssump}.  We  also want to satisfy assumption \eqref{addAssump}.  
%To facilitate the fast transform, it is assumed in this section and the next that $n$ is power of $2$.  
To facilitate the fast transform, $n$ must be power of $2$.

@@

\subsection{Shift Invariant Kernels}
The covariance functions $C$ that match integration lattice node sets have the form
\begin{align} \label{eq:shInv}
C(\vt,\vx) = K(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift invariant kernel} because shifting both arguments of the covariance function by the same amount leaves the value unchanged.   By a proper scaling of the kernel $K$ it follows that assumption \eqref{addAssump} is satisfied. Of course, $K$ must also be of the form that ensures that $C$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.

>>>>

\newcommand{\hatvy}{\hat{\bm{y}}}
\newcommand{\haty}{\hat{y}}
\newcommand{\tvy}{\tilde{\bm{y}}}
\newcommand{\ty}{\tilde{y}}
\newcommand{\vzero}{\bm{0}}
\newcommand{\vone}{\bm{1}}
\newcommand{\tvone}{\tilde{\bm{1}}}
\newcommand{\mA}{\mathsf{A}}
\newcommand{\mC}{\mathsf{C}}
\newcommand{\mP}{\mathsf{P}}
\newcommand{\mK}{\mathsf{K}}
\newcommand{\rmC}{\mathring{\mathsf{C}}}
\newcommand{\mCtheta}{{\mathsf{C}_{\vtheta}}}
%\newcommand{\mCthetaInv}{{\mathsf{C}^{-1}_{\vtheta}}}
%\newcommand{\mCthetaMLE}{{\mathsf{C}_{\vthetaMLE}}}
%\newcommand{\mCthetaInvMLE}{{\mathsf{C}^{-1}_{\vthetaMLE}}}
\newcommand{\mCInv}{\mathsf{C}^{-1}}
\newcommand{\cov}{{\textup{cov}}}
\newcommand{\var}{{\textup{var}}}
\newcommand{\opt}{{\textup{opt}}}

@@

\allowdisplaybreaks
\begin{document}
\setlength\abovedisplayskip{1pt}
\setlength{\belowdisplayskip}{1pt}

@@

The primary contribution of this article is to demonstrate how the choice of a family of covariance kernels that match the low discrepancy sampling nodes facilitates fast computation of the cubature and the data-driven stopping criterion.  Our Bayesian cubature requires a computational cost of
\begin{equation} \label{eqn:OuralgoCost}
    \Order\bigl(n \$(f) + N_{\textup{opt}}[n\$(C) + n \log(n)] \bigr),
\end{equation} 
where $\$(f)$ is the cost of one integrand value, $\$(C)$ is the cost of a single covariance kernel value,  $\Order(n \log(n))$ is the cost of a fast Fourier transform, and $N_{\textup{opt}}$ is an upper bound on the number of optimization steps required to choose the hyperparameters. If function evaluation is expensive, such as coming from computationally intensive simulation, or $\$(f) = \Order(d)$ for large $d$, then $\$(f)$ might be similar in magnitude to $N_{\textup{opt}} \log(n)$ in practice.  Typically, $\$(C) = \Order(d)$.  Note that the $\Order(n \log(n))$ contribution is $d$ independent.

By contrast to our fast algorithm, the typical computational cost for Bayesian cubature is
\begin{equation} \label{eqn:TheiralgoCost}
    \Order\bigl(n \$(f) + N_{\textup{opt}}[n^2\$(C) + n^3] \bigr),
\end{equation} 
which is explained in Section \ref{sec:bayes_cubature_algo}. Note that the work beyond evaluating the integrand in \eqref{eqn:TheiralgoCost} is much larger than that in \eqref{eqn:OuralgoCost}.

@@

Section \ref{sec:BC} explains the Bayesian approach to estimate the posterior cubature error and defines our automatic Bayesian cubature. Although much of this material is known, it is included for completeness.  We end Section \ref{sec:BC}  by demonstrating why Bayesian cubature is typically computationally expensive.
Section \ref{sec:fast_BC}  introduces the concept of covariance kernels that match the nodes and expedite the computations required by our automatic Bayesian cubature. 
Section \ref{sec:shift_invariant_kernel} implements this concept for shift-invariant kernels and rank-1 lattice nodes.  It also describes how to avoid cancellation error for kernels of product form.  Numerical examples are provided in Section \ref{sec:NumExp} to demonstrate our new algorithm.  We conclude with a brief discussion.

@@

\begin{theorem} \label{thm:param} There are at least three approaches to estimating or integrating out the hyperparameters defining the Gaussian process from which the integrand is drawn: empirical Bayes, full Bayes, and generalized cross validation.  Under these three approaches we have the following:
%\begin{allowdisplaybreaks}
\begin{align}
    \label{eqn_m_MLE}
m_\MLE &= \frac{\vone^T \mCInv_\vtheta \vy }{ \vone^T \mCInv_\vtheta \vone}, \qquad
m_{\textup{GCV}} = \frac{\vone^T \mC_\vtheta^{-2} \vy}{\vone^T \mC_\vtheta^{-2} \vone}, \\
\label{eqn_s2_MLE}
s^2_{\MLE} 
&= 
\frac{1}{n}
\vy^T 
\left[ \mCInv_\vtheta - 
\frac{ \mCInv_\vtheta \vone \vone^T \mCInv_\vtheta }{\vone^T\mCInv_\vtheta \vone}
\right] \vy, \\
\nonumber
\hsigma_{\textup{full}}^2 
& = \frac{1}{n-1}
\vy^T\left[ \mC_\vtheta^{-1} 
- \frac{ \mC_\vtheta^{-1} \vone\vone^T \mC_\vtheta^{-1}}{\vone^T \mC_\vtheta^{-1} \vone}  \right]\vy
\\ 
& \times  \left[\frac{(1 - \vc_{\vtheta}^T \mC_\vtheta^{-1} \vone)^2}{\vone^T \mC_\vtheta^{-1} \vone} + (c_{0\vtheta}  -\vc_{\vtheta} ^T \mC_\vtheta^{-1} \vc_{\vtheta}) \right], \\
\nonumber
 s^2_{\textup{GCV}} & = \vy^T \left[\mC_\vtheta^{-2} - \frac{\mC_\vtheta^{-2} \vone \vone^T \mC_\vtheta^{-2}}{\vone^T \mC_\vtheta^{-2} \vone}  \right] \vy  \left[ \trace(\mC_\vtheta^{-1}) \right]^{-1}, \\
\nonumber
\vtheta_\MLE
&= \argmin_{\vtheta} \biggl \{
\log\left(\vy^T 
\left[ \mC_\vtheta^{-1} - 
\frac{ \mCInv_\vtheta \vone \vone^T \mCInv_\vtheta }{\vone^T\mCInv_\vtheta \vone}
\right] \vy 
\right)  \\
\label{eqn:thetaMLE}
 & \qquad +  \frac{1}{n} \log(\det(\mC_\vtheta))
\biggr \}, \\
\nonumber
\vtheta_{\textup{GCV}} &= \argmin_\vtheta \biggl\{\log \left(  \vy^T \left[\mC^{-2}_\vtheta - \frac{\mC^{-2}_\vtheta \vone \vone^T \mC^{-2}_\vtheta}{\vone^T \mC^{-2}_\vtheta \vone}  \right] \vy \right)  \\
\label{vthetaGCV}
& - \log \left ( \trace(\mC^{-2}_\vtheta) \right ) \biggr\}, \\
\hmu_\MLE  &= \hmu_\full =
\left(
\frac{ (1 - \vone^T  \mCInv_\vtheta\vc_{\vtheta} )  \vone }{ \vone^T \mCInv_\vtheta \vone}   +  \vc_{\vtheta} 
\right)^T  \mCInv_\vtheta \vy, \\
\label{eqn:muCV}
\widehat{\mu}_{\textup{GCV}}
& = \left(\frac{(1 - \vone^T  \mC_\vtheta^{-1}\vc_{\vtheta}) \mC_\vtheta^{-1} \vone}{\vone^T \mC_\vtheta^{-2} \vone} + \vc_{\vtheta} \right)^T \mC_\vtheta^{-1} \vy, \\ 
\nonumber
\err_{\mathsf{x}} & = 2.58 s_{\mathsf{x}} \sqrt{c_{0\vtheta} - \vc_{\vtheta}^T\mC_\vtheta^{-1}\vc_{\vtheta} }, \\
& \hspace{3.5cm} \mathsf{x} \in \{\MLE, \GCV\},  \label{eqn:err_MLEGCV}\\ 
\err_{\textup{full}} 
& = t_{n-1,0.995} \hsigma_{\textup{full}} > \err_\MLE, \label{FJH:eq:errFull}
\end{align}
%\end{allowdisplaybreaks}
\begin{multline}
\label{eqn_prob_CI}
\mathbb{P}_f \left[
|\mu-\hmu_{\mathsf{x}}| \leq \err_{\mathsf{x}} \right]  = 99\%, \\ \mathsf{x} \in \{\MLE, \full, \GCV\}.
\end{multline}
Here $t_{n-1,0.995}$ denotes the $99.5$ percentile of a standard Student's $t$-distribution with $n-1$ degrees of freedom.
\end{theorem}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proof for Empirical Bayes}  \label{sec:MLE}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The empirical Bayes approach estimates the parameters, $m$, $s$, and $\vtheta$ via maximum likelihood estimation (MLE).  The log-likelihood function of the parameters given the function data $\vy$ is:
\begin{multline*}
l(s,m,\vtheta | \vy)
= -\frac{1}{2} s^{-2} (\vy-m\vone)^T\mCInv_\vtheta(\vy-m\vone) 
\\
 - \frac{1}{2} \log(\det\, \mC_\vtheta) - \frac{n}{2} \log(s^2) + \text{constants.}
\end{multline*}
Maximizing the log-likelihood first with respect to $m$ and then with respect to $s$ yields the values given in Theorem \ref{thm:param}.
To obtain $\vtheta_\MLE$, we substitute $m_\MLE$ and $s_\MLE$ into $l(s,m,\vtheta | \vy)$, which leads directly to the optimization problem in \eqref{eqn:thetaMLE}.

The empirical Bayes estimate of $\vtheta$ balances minimizing the covariance scale factor, $s^2_{\MLE}$, against minimizing  $\det(\mC_\vtheta)$. 
Under these estimates of the parameters, the cubature \eqref{eqn:BayesCub} and the credible interval \eqref{eqn_prob_confidence_interval} are explicitly written as in Theorem \ref{thm:param}.
The quantities $c_{0\vtheta}$, $\vc_\vtheta$, and $\mC_\vtheta$ are assumed implicitly to be based on $\vtheta = \vtheta_\MLE$.

@@

        \State\label{LoopStart}Generate $\{ \vx_i\}_{i=n' + 1}^{n}$ and sample $\{f(\vx_i)\}_{i=n'+1}^{n}$
        \State Compute $\vtheta$ by \eqref{eqn:thetaMLE} or \eqref{vthetaGCV}
        \State Compute $\err_{\textup{CI}}$  according to \eqref{eqn:err_MLEGCV} or \eqref{FJH:eq:errFull}

@@

        \State Update sample size to compute $\hmu$, $n \gets n'$
        \State Compute $\hmu$, the approximate integral,   according to \eqref{eqn:cubMLE} or \eqref{eqn:muCV}
      \State \Return $\hmu$, $n$,  and $\err_{\textup{CI}}$
  \end{algorithmic}
\end{algorithm}

As described above, the computational cost of Algorithm \ref{algorithm1} is the sum of the following:
\begin{itemize}
	\item $\Order\bigl(n\$(f) \bigr)$ for the function data, where $\$(f)$ is the computational cost of a single $f(\vx)$; $\$(f)$ may be large if it is the result of an expensive simulation; $\$(f)$ is typically proportional to $d$;

	\item $\Order\bigl(N_{\opt} n^2 \$(C_\vtheta) \bigr)$ for the evaluation of the matrix $\mC_{\vtheta}$, $N_\opt$ is the number of optimization steps required, and  $\$(C_\vtheta)$ is the computational cost of a single $C_\vtheta(\vt,\vx)$; $\$(C)$ is typically proportional to $d$; and

	\item $\Order\bigl(N_{\opt} n^3 \bigr)$ for the matrix inversions and determinant calculations.

\end{itemize}
As we see in the example in the next section, this cost increases quickly as the $n$ required to meet the error tolerance increases.  This motivates the fast Bayesian cubature algorithm presented in Section \ref{sec:fast_BC}.

@@

Genz \cite{Gen93} introduced a variable transformation to transform \eqref{eqn:GaussDef} into an integral on the unit cube.  Let $\mSigma= \mL \mL^T$ be the Cholesky decomposition where $\mL = (l_{jk})_{j,k=1}^d$ is a lower triangular matrix.  Iteratively define
\allowdisplaybreaks
\begin{align*}
\alpha_1& = \Phi(a_1), \qquad \beta_1 = \Phi(b_1)
\\
\alpha_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
a_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\\
\beta_j&(x_1,...,x_{j-1}) = 
\\
&\; \Phi
\left(
\frac{1}{l_{jj}} 
\left(
b_j - \sum_{k=1}^{j-1} l_{jk} \Phi^{-1}(\alpha_k + x_k(\beta_k-\alpha_k))
\right)
\right), 
\\
&\hspace{5cm} j=2,...,d,
\end{align*}
\begin{align}
\label{fGenzdef}
f_{\text{Genz}}(\vx) = \prod_{j=1}^d [\beta_j(\vx) - \alpha_j(\vx)].
\end{align}
where $\Phi$ is the cumulative standard normal distribution function.  Then, $\mu = \int_{[0,1]^{d'-1}} f_{\text{Genz}}(\vx) \, \dvx$. As we can see, this approach transforms $d'$ integral into a $d=d'-1$ integral.

@@

We use the following parameter values in the simulation: 
\begin{align*}
d' = 3, \quad \va = \begin{pmatrix}
-6 \\ -2 \\ -2
\end{pmatrix}, \quad 
\vb = \begin{pmatrix}
5 \\ 2 \\ 1
\end{pmatrix} , \quad 
\mL = \begin{pmatrix}
4 & 1 & 1 \\ 0 & 1 & 0.5 \\ 0 & 0 & 0.25
\end{pmatrix}.
\end{align*}
The node sets are randomly scrambled Sobol' points \cite{DicEtal14a,DicPil10a}.  The results are for 400 randomly chosen $\varepsilon$ in the interval $[10^{-5}, 10^{-2}]$ as shown in \figref{fig:MVN_Metern_d2b2}. In each run, the nodes are randomly scrambled.  We  observe the algorithm meets the error criterion 95\% of the time.
%with Intel i7 3630QM and 16GB RAM memory
%On our test computer, 
As shown in \figref{fig:MVN_Metern_d2b2}, computation time increases rapidly with $n$. 
The  maximum likelihood estimation of $\vtheta$, which requires repeated evaluation of the objective function, is the most time consuming of all. It takes tens of seconds to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$.   In contrast, this example in Section \ref{sec:NumExp} take less than a hundredth of a second to compute $\hmu_n$ once with $\varepsilon = 10^{-5}$ using our new algorithm. Not only is the Bayesian cubature with the Mat\'ern kernel slow, but also $\mC_\vtheta$ becomes highly ill-conditioned as $n$ increases.
So, Algorithm \ref{algorithm1}  in its current form is impractical when $n$ must be large. 
%\FJHNote{Jags, please check this re-write} %\JRNote{looks good}

@@

The generic automatic Bayesian cubature algorithm described in the previous section requires $\Order\bigl(n \$(f) + \break N_{\opt} [n^2 \$(C_\vtheta)  + n^3]\bigr)$ operations to estimate $\vtheta$, compute the credible interval width, and compute the cubature. Now we explain how to speed up the calculations. A key is to choose covariance kernels that match the nodes, $\{\vx_i\}_{i=1}^n$, so that the vector-matrix operations required by Bayesian cubature can be accomplished using fast transforms at a cost of $\Order\bigl(n \$(f) + N_{\opt} [n \$(C_\vtheta)  + n \log(n)] \bigr)$.

@@

Under assumptions \eqref{fastcompAssump} the eigenvalues may be identified as the fast transform of the first column of $\mC$:
\begin{align}
\nonumber
\vlambda 
& = \begin{pmatrix}
\lambda_1 \\ \vdots \\ \lambda_n
\end{pmatrix} = \mLambda \vone = \mLambda \vv_1^* 
= \underbrace{\left( \frac 1n \mV^H  \mV \right) }_{\mathsf{I}} \mLambda \vv_1^* \\
&= \mV^H \left( \frac 1n \mV \mLambda \vv_1^* \right)
= \mV^H \vC_1 =  \widetilde{\vC}_1,
\label{eqn:fast_transform_to_eigvalues}
\end{align}
where $\mathsf{I}$ is the identity matrix and $\vv_1^*$ is the complex conjugate of the first row of $\mV$.
Also note that the fast transform of $\vone$ has a simple form
\begin{align*} 
\widetilde{\vone}
& = \mV^H \vone = \mV^H \vV_1 = 
\left(n, 0, \dots, 0 \right)^T.
% \begin{pmatrix}n  \\ 0 \\ \vdots \\ 0 \end{pmatrix}.
\label{eqn:fast_transform_one}
\end{align*}

@@

The preceding sections lay out an automatic Bayesian cubature algorithm whose computational cost is drastically reduced.  However, this algorithm relies on covariance kernel functions, $C_\vtheta$ and node sets, $\{\vx_i\}_{i=1}^n$ that satisfy assumptions \eqref{fastcompAssump}.  We  also want to satisfy assumption \eqref{addAssump}.  
%To facilitate the fast transform, it is assumed in this section and the next that $n$ is power of $2$.  
To facilitate the fast transform, $n$ must be power of $2$.

@@

\subsection{Shift Invariant Kernels}
The covariance functions $C_\vtheta$ that match integration lattice node sets have the form
\begin{align} \label{eq:shInv}
C_\vtheta(\vt,\vx) = K_\vtheta(\vt - \vx \bmod \vone).
\end{align}
This is called a \emph{shift-invariant kernel} because shifting both arguments of the covariance function by the same amount leaves the value unchanged.   By a proper scaling of the kernel $K$ it follows that assumption \eqref{addAssump} is satisfied. Of course, $K_\vtheta$ must also be of the form that ensures that $C_\vtheta$ is symmetric and positive definite, as assumed in \eqref{FJH:eq:CondPosDef}.
